---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---



```{r}
library(data.table)
```

I have a problem with a join that should produce extra rows in data table, I have a subresult which is this one:

```{r generateSubresult}
subresult <- structure(list(cell = c(1447386L, 1447386L, 1447387L), species = c("Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana"), Habitat = c("WetRich", "DryRich", "DryRich")), class = c("data.table", 
"data.frame"), row.names = c(NA, -3L))
```


this looks like this:

```{r, echo = F}
knitr::kable(subresult)
```

Here we have a species that can inhabit in 2 different habitats in one cell, and 1 in another, all of which are available in this cell, now those habitats can become different landuses, so we have this other dataset

```{r}
Feasible_Landuses <- structure(list(species = c("Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana"), Landuse = c("ForestDryRich", 
"ForestDryPoor", "ForestWetRich", "ForestWetPoor", "OpenDryRich", 
"OpenDryPoor", "OpenWetPoor"), Habitat = c("DryRich", "DryPoor", 
"WetRich", "WetPoor", "DryRich", "DryPoor", "WetPoor")), row.names = c(NA, 
-7L), class = c("data.table", "data.frame"), sorted = "species")
```

which can be seen here

```{r, echo = F}
knitr::kable(Feasible_Landuses)
```

Now the goal would be to make a join resulting in this:

```{r}
Final_Result <- structure(list(cell = c(1447386L, 1447386L, 1447386L, 1447387L, 1447387L), species = c("Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana"), Landuse = c("ForestDryRich", "ForestWetRich", 
"OpenDryRich", "ForestDryRich","OpenDryRich"), Habitat = c("DryRich", "WetRich", "DryRich", "DryRich", "DryRich")), row.names = c(NA, 
-5L), class = c("data.table", "data.frame"))
```

Which should look like this:

```{r, echo = F}
knitr::kable(Final_Result)
```

When I try this this works fine

```{r}
Final_Result <- subresult[Feasible_Landuses, on = .(Habitat, species), nomatch = 0]
```

as can be seen here

```{r, echo=FALSE}
knitr::kable(Final_Result)
```


However on a larget dataset I get this error:

```{r, eval = F}
Error in vecseq(f__, len__, if (allow.cartesian || notjoin || !anyDuplicated(f__,  : 
  Join results in 19648 rows; more than 11143 = nrow(x)+nrow(i). Check for duplicate key values in i each of which join to the same group in x over and over again. If that's ok, try by=.EACHI to run j for each group to avoid the large allocation. If you are sure you wish to proceed, rerun with allow.cartesian=TRUE. Otherwise, please search for this error message in the FAQ, Wiki, Stack Overflow and data.table issue tracker for advice.
```

## Now with a big dataset

this dataset is available at [this repo](https://github.com/derek-corcoran-barrios/DataTableJoinQuestion)

But I will add the code so that I read it from there:



## Session info

