---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---
# Summary:

I am encountering a perplexing issue with a data.table join operation in R, where I intend to combine information from two datasets based on specific habitat and species criteria. The join is expected to yield additional rows, but on a large dataset, it results in a "Join results in more rows" error. To illustrate the problem, I've provided a minimal example with sample datasets. Initially, a smaller dataset demonstrates successful execution of the join, but the same operation on a larger dataset triggers an error. I've included the relevant code snippets, sample data, and error messages. Seeking guidance on resolving this issue, particularly when dealing with duplicate key values and large datasets, I am reaching out for insights and potential solutions.

# This is a small dataset, here it works
```{r}
library(data.table)
```

I have a problem with a join that should produce extra rows in data table, I have a subresult which is this one:

```{r generateSubresult}
subresult <- structure(list(cell = c(1447386L, 1447386L, 1447387L), species = c("Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana"), Habitat = c("WetRich", "DryRich", "DryRich")), class = c("data.table", 
"data.frame"), row.names = c(NA, -3L))
```


this looks like this:

```{r, echo = F}
knitr::kable(subresult)
```

Here we have a species that can inhabit in 2 different habitats in one cell, and 1 in another, all of which are available in this cell, now those habitats can become different landuses, so we have this other dataset

```{r}
Feasible_Landuses <- structure(list(species = c("Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana"), Landuse = c("ForestDryRich", 
"ForestDryPoor", "ForestWetRich", "ForestWetPoor", "OpenDryRich", 
"OpenDryPoor", "OpenWetPoor"), Habitat = c("DryRich", "DryPoor", 
"WetRich", "WetPoor", "DryRich", "DryPoor", "WetPoor")), row.names = c(NA, 
-7L), class = c("data.table", "data.frame"), sorted = "species")
```

which can be seen here

```{r, echo = F}
knitr::kable(Feasible_Landuses)
```

Now the goal would be to make a join resulting in this:

```{r}
Final_Result <- structure(list(cell = c(1447386L, 1447386L, 1447386L, 1447387L, 1447387L), species = c("Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana", "Abies nordmanniana", 
"Abies nordmanniana"), Landuse = c("ForestDryRich", "ForestWetRich", 
"OpenDryRich", "ForestDryRich","OpenDryRich"), Habitat = c("DryRich", "WetRich", "DryRich", "DryRich", "DryRich")), row.names = c(NA, 
-5L), class = c("data.table", "data.frame"))
```

Which should look like this:

```{r, echo = F}
knitr::kable(Final_Result)
```

When I try this this works fine

```{r}
Final_Result <- subresult[Feasible_Landuses, on = .(Habitat, species), nomatch = 0]
```

as can be seen here

```{r, echo=FALSE}
knitr::kable(Final_Result)
```

# Error on large dataset

However on a larget dataset I get this error:

```{r, eval = F}
Error in vecseq(f__, len__, if (allow.cartesian || notjoin || !anyDuplicated(f__,  : 
  Join results in 19648 rows; more than 11143 = nrow(x)+nrow(i). Check for duplicate key values in i each of which join to the same group in x over and over again. If that's ok, try by=.EACHI to run j for each group to avoid the large allocation. If you are sure you wish to proceed, rerun with allow.cartesian=TRUE. Otherwise, please search for this error message in the FAQ, Wiki, Stack Overflow and data.table issue tracker for advice.
```

## Now with a big dataset

this dataset is available at [this repo](https://github.com/derek-corcoran-barrios/DataTableJoinQuestion)

But I will add the code so that I read it from there:

```{r}
result <- data.table::fread("https://raw.githubusercontent.com/derek-corcoran-barrios/DataTableJoinQuestion/master/result.csv")
```

you can see the structure here

```{r}
str(result)
```

and now the feasible landuses

```{r}
Feasible_Landuses <- data.table::fread("https://raw.githubusercontent.com/derek-corcoran-barrios/DataTableJoinQuestion/master/Feasible_Landuses.csv")
```

Again we check the structure

```{r}
str(Feasible_Landuses)
```

and now we finally check the join:

```{r}
Final_Result <- result[Feasible_Landuses, on = .(Habitat, species), nomatch = 0]
```

Why this one gets an error and this first one does not?

## Session info

